---
title: "BIST 5092 Phase III Project"
output: pdf_document
author: Diezhang Wu
---
```{r}
library(dplyr)
library(tidyr)
library(stringr)
```

### import ADPA and ADSL datasets, select measurements at visit 6
```{r}
adpa <- read.csv("../data_phase 3/ADPA.csv")
pasi <- adpa %>% filter(PARAMCD == "PASI")
pasi6 <- adpa %>% filter(PARAMCD == "PASI", AVISIT == "VISIT 6")
adsl <- read.csv("../data_phase 3/ADSL.csv")
adae <- read.csv("../data_phase 3/ADAE.csv")
```

### perform logistic regression to assess the treatment effect on PASI75 adjusted by sex with the full dataset
```{r}
dat_full <- merge(adsl, pasi6, by = "SUBJID")
summary(glm(PCHGCA1N ~ SEX + as.factor(TRTPN), data = dat_full, family = binomial))
```

## MCAR
### create functions to generate dataset with certain percent missing
```{r}
generate_missing <- function(data, percent){
  id <- sample(1:nrow(data), round(nrow(data)*percent), replace = FALSE)
  data_par1 <- data[id,]
  data_par1$AVAL <- NA
  data_par2 <- data[-id,]
  data_final <- rbind(data_par1, data_par2) 
  data_final$PCHGCA1N[which(is.na(data_final$AVAL))] <- NA
  
  return(data_final)
}
```

### 1. generate dataset with 10% missing and perform logistic regression 
```{r}
set.seed(1234)
pasi6_final1 <- generate_missing(pasi6, 0.1)
dat1 <- merge(adsl, pasi6_final1, by = "SUBJID")
summary(glm(PCHGCA1N ~ SEX + as.factor(TRTPN), data = dat1, family = binomial))
```

### 2. Impute the missingness and reanalyze the data
```{r}
dat1_im <- dat1
dat1_im$PCHGCA1N[which(is.na(dat1_im$AVAL))] <- 0
summary(glm(PCHGCA1N ~ SEX + as.factor(TRTPN), data = dat1_im, family = binomial))
```


### 3. repeat 1 & 2 with 20% missingness
```{r}
## with 20% missingness
set.seed(1234)
pasi6_final2 <- generate_missing(pasi6, 0.2)
dat2 <- merge(adsl, pasi6_final2, by = "SUBJID")
summary(glm(PCHGCA1N ~ SEX + as.factor(TRTPN), data = dat2, family = binomial))

## with 20% missingness imputed
dat2_im <- dat2
dat2_im$PCHGCA1N[which(is.na(dat2_im$AVAL))] <- 0
summary(glm(PCHGCA1N ~ SEX + as.factor(TRTPN), data = dat2_im, family = binomial))
```

### repeat 1 & 2 with 30% missingness
```{r}
## with 30% missingness
set.seed(1234)
pasi6_final3 <- generate_missing(pasi6, 0.3)
dat3 <- merge(adsl, pasi6_final3, by = "SUBJID")
summary(glm(PCHGCA1N ~ SEX + as.factor(TRTPN), data = dat3, family = binomial))

## with 30% missingness imputed
dat3_im <- dat3
dat3_im$PCHGCA1N[which(is.na(dat3_im$AVAL))] <- 0
summary(glm(PCHGCA1N ~ SEX + as.factor(TRTPN), data = dat3_im, family = binomial))
```

## MAR
```{r}
# select subject id, visit number, percent change from baseline at each visit
# reshape the dataset into wide format
pasi_inter <- pasi %>% 
  select(SUBJID, AVISIT, PCHG) %>% 
  spread(AVISIT, PCHG)

# generate missing data conditioned on previous visits
# in miss4, miss5 and miss6, 0 is missing, 1 is not missing
set.seed(1234)
for (i in 1:nrow(pasi_inter)){
  if (pasi_inter$`VISIT 3`[i] < 10){
    pasi_inter$miss4[i] <- rbinom(1, 1, 0.7)
  } else{ pasi_inter$miss4[i] <- rbinom(1, 1, 0.95)}
}

for (i in 1:nrow(pasi_inter)){
  if (pasi_inter$miss4[i] == 0){
    pasi_inter$miss5[i] = 0
  } else if (pasi_inter$`VISIT 4`[i] < 10){
    pasi_inter$miss5[i] <- rbinom(1, 1, 0.7)
  } else{ pasi_inter$miss5[i] <- rbinom(1, 1, 0.95)}
} 

for (i in 1:nrow(pasi_inter)){
  if (pasi_inter$miss4[i] == 0){
    pasi_inter$miss6[i] = 0
  } else if (pasi_inter$`VISIT 5`[i] < 10){
    pasi_inter$miss6[i] <- rbinom(1, 1, 0.7)
  } else{ pasi_inter$miss6[i] <- rbinom(1, 1, 0.95)}
}

pasi6_mar <- pasi6
pasi6_mar$PCHGCA1N[which(pasi_inter$miss6 == 0)] <- NA # PCHGCA1N contains visit 6 results with missing values
pasi6_mar$im6 <- pasi_inter$miss6 # im6 contains the imputed visit 6 results

dat_mar <- merge(adsl, pasi6_mar, by = "SUBJID")
```


```{r}
# logistic regression on PASI75 with missing values excluded
summary(glm(PCHGCA1N ~ SEX + as.factor(TRTPN), data = dat_mar, family = binomial))

# logistic regression on PASI75 with missing data imputed as 0
summary(glm(im6 ~ SEX + as.factor(TRTPN), data = dat_mar, family = binomial))
```

## MNAR
```{r}
# sum the duration of each adverse event to generate total days of adverse event the patient had during the trial
adae$duration <- adae$AEENDY - adae$AESTDY
#adduration <- adae %>% group_by(USUBJID) %>% summarise(duration = sum(duration))

# create severity score for each patient with AE using AESER, AESEV, and AEREL
# AESER_rate = 1 for Y, AESEV_rate : 1 for mild, 2 for moderate, 3 for severe, AEREL_rate = 1 for Y
adae$AESER_rate <- ifelse(adae$AESER == "Y", 1, 0)
adae$AESEV_rate <- ifelse(adae$AESEV == "MILD", 1, ifelse(adae$AESEV == "MODERATE", 2, 3))
adae$AEREL_rate <- ifelse(adae$AEREL == "Y", 1, 0)

# severity score score per event = AEREL_rate * 5 + AESER_rate * 3 + AESEV_rate * 2
adae$serscore <- adae$AEREL_rate * 5 + adae$AESER_rate * 3 + adae$AESEV_rate * 2

# toxic score = severity score * duration
adae$toxscore <- adae$serscore * adae$duration 

# compute toxicity scale for each subject by sum the toxic score from each adverse event
toxscale <- adae %>% group_by(USUBJID) %>% summarise(toxscore = sum(toxscore))
# check the distribution of the toxic scale
summary(toxscale$toxscore)
hist(toxscale$toxscore)

hist(log(toxscale$toxscore))

# assume the probability of dropping out is positively related to the toxicity scale and has an uniform distribution
# the probability then can be simply determined by dividing the toxicity score with a common denominator
toxscale$prob <- toxscale$toxscore/2000
summary(toxscale$prob)
hist(toxscale$prob)

toxscale$problog <- log(toxscale$toxscore)/10
summary(toxscale$problog)
hist(toxscale$problog)

# impute the missingness of the subjects with adverse events with their probability of dropping out
aemissing <- toxscale
set.seed(10)
for (i in 1:nrow(aemissing)){aemissing$miss[i] <- rbinom(1, 1, 1 - aemissing$prob[i])}
for (i in 1:nrow(aemissing)){aemissing$misslog[i] <- rbinom(1, 1, 1 - aemissing$problog[i])}

aemissing$SUBJID <- as.numeric(substr(aemissing$USUBJID, nchar(aemissing$USUBJID) - 4 + 1, nchar(aemissing$USUBJID)))

# join the aemissing table to pasi6 and set visit6 results to be missing based on results from the previous step (1 = not missing, 0 = missing)
dat_mnar <- left_join(pasi6, aemissing, by = "SUBJID")
dat_mnar$PCHGCA1N[which(dat_mnar$miss == 0)] <- NA
dat_mnar$im6 <- dat_mnar$PCHGCA1N
dat_mnar$im6[which(is.na(dat_mnar$PCHGCA1N))] <- 0
dat_mnar <- merge(dat_mnar, adsl, by = "SUBJID")

# perform logistic regression on pasi75 with incomplete dataset and imputed dataset
summary(glm(PCHGCA1N ~ SEX + as.factor(TRTPN), data = dat_mnar, family = binomial))
summary(glm(im6 ~ SEX + as.factor(TRTPN), data = dat_mnar, family = binomial))
```

